---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

The data I chose to analyze is Seattle Public Library's checkout data for items that have been checked out at least 10 times a month from 2017-2023. This data offered insight into what items were the most popular to library users, as well as reveal the trend on the popularity of those items for the past six years. I wanted to see if worldwide events like the pandemic and BLM movement influenced the items that were most frequently checked out in libraries.  

### Summary Information

```{r include=FALSE, paged.print=TRUE}
source("summary.R")
```
From 2017 to 2023, there were `r num_unique_items` unique items that were checked out more 10 times each month from the Seattle Public Libraries. Of those `r num_unique_items` unique items, `r physical_checkouts` were physical items, and `r digital_checkouts` were digital items. The month with the highest number of checkouts was `r month_most_checkouts$date` (ignoring the day value), where there was a total of `r month_most_checkouts$totalCheckouts` checkouts for items that were checked out 10+ times monthly. On the other hand, the month with the least checkouts for these items was `r month_least_checkouts$date` (ignoring the day value), where there was a total of `r month_least_checkouts$totalCheckouts` checkouts. 

During the time frame of 2017 to 2023, the most checked out item in a singular month was the `r highest_checkout$MaterialType`, `r highest_checkout$Title` by `r highest_checkout$Creator`. It was checked out `r highest_checkout$Checkouts` times in the month of `r highest_checkout$date` (ignoring the day value). I found this data very interesting because the time coincided with the BLM movement that took place worldwide, so it was cool to see how the library checkout data reflected what was going on in our society. 

### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### Chart 1: Average Checkout Rates over Time

The first chart shows the average number of checkouts of a single item in a month for items that have been checked out for 10+ times monthly. I wanted to look at the average checkout times for popular library items over time to see when there were higher checkout rates, and to visually show that trend I displayed the data through a line chart. Based on the line chart, I found that June of 2020 had the highest average checkout times for single items, with an average of 28.9 checkouts each month for one item. This was an interesting finding because this was also the time when the pandemic was in its peak and people were quarantined indefinitely, so it made sense to see an increase in the average checkout rates around that time period.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("chart1.R")
average
```

### Chart 2: Types of Items Checked Out over Time

The second chart showed the total number of checkouts for items that have been checked out 10+ times monthly over the past six years (2017-2023) and their composition of whether the items were physical or digital materials. Based on this bar graph, We can see that generally there are a total of around 3 million total checkouts for these items each year, which the 2023 data being a bit of an outlier since the year just started. However, I found it super interesting that the composition of those checkouts based on material type had a drastic change starting in 2020. The Covid-19 pandemic shifted much of our world's operations an online model, and that is also reflected in this checkout data. Before 2020, the majority of checkouts were physical items, but starting in 2020, the majority of checkouts were digital items. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("chart2.R")
barplot
```

### Chart 3: Most Chekced Out Material Types

The last chart I analyzed was a piechart that showed the composition of all the checkouts from 2017-2023 based on their material type. I chose to examine this data because I think libraries are typically associated with borrowing books, but libraries actually offer a much wider medium of materials. The Seattle Public Library offers 32 types of items to checkout, and this piechart shows the top 10 most checked out material types. From the chart, we can see that the four most checked out material type are audiobooks, books, eBooks, and videodiscs. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("chart3.R")
piechart
```


